Lactate transport inhibition therapeutically reprograms fibroblast

### 2024-03-20 manuscript/manuscript.qmd
add EMS acknowledgement
update authors

authors:
- name: David R. Ziehr
  affiliations: 
  - ref: bwh
  - ref: mgh
  - ref: hms
- name: Fei Li
  affiliations:
  - ref: bwh
  - ref: hms
- name: Nathan M. Krah
  affiliations:
  - ref: utah_gen
- name: Kevin J. Leahy 
  affiliations:
  - ref: bwh
- name: K. Mark Parnell
  affiliations:
  - ref: vettore
- name: Christelle Guillermier
  affiliations:
  - ref: bwh
  - ref: hms
- name: Jack Varon
  affiliations: 
  - ref: bwh
  - ref: hms
- name: Rebecca M. Baron
  affiliations:
  - ref: bwh
  - ref: hms
- name: Bradley A. Maron
  affiliations:
  - ref: umd
  - ref: ihc
- name: Nancy J. Philp
  affiliations:
  - ref: tj
- name: Lida P. Hariri
  affiliations: 
  - ref: hms
  - ref: mgh_path
- name: Edy Y. Kim
  affiliations:
  - ref: bwh
  - ref: hms
- name: Matthew L. Steinhauser
  affiliations:
  - ref: pitt
  - ref: upmc
- name: Rachel S. Knipe
  affiliations:
  - ref: mgh
  - ref: hms
- name: Jared Rutter
  affiliations:
  - ref: utah_biochem
  - ref: utah_hhmi
- name: William M. Oldham
  affiliations:
  - ref: bwh
  - ref: hms
  email: woldham@bwh.harvard.edu
  corresponding: true

affiliations:
- id: bwh
  name: Department of Medicine, Brigham and Women's Hospital, Boston, MA
- id: mgh
  name: Department of Medicine, Massachusetts General Hospital, Boston, MA
- id: hms
  name: Department of Medicine, Harvard Medical School, Boston, MA
- id: utah_gen
  name: Department of Human Genetics, University of Utah, Salt Lake City, UT
- id: vettore
  name: Vettore Biosciences, San Francisco, CA
- id: umd
  name: Department of Medicine, University of Maryland School of Medicine, Baltimore, MA
- id: ihc
  name: University of Maryland Institute for Health Computing, Bethesda, MD
- id: tj
  name: Department of Pathology, Anatomy & Cell Biology, Thomas Jefferson University, Philadelphia, PA
- id: mgh_path
  name: Department of Pathology, Massachusetts General Hospital, Boston, MA
- id: pitt
  name: Aging Institute, University of Pittsburgh, Pittsburgh, PA
- id: upmc
  name: UPMC Heart and Vascular Institute, UPMC Presbyterian, Pittsburgh, PA
- id: utah_biochem
  name: Department of Biochemistry, University of Utah, Salt Lake City, UT
- id: utah_hhmi
  name: Howard Hughes Medical Institute, University of Utah School of Medicine, Salt Lake City, UT

@@

Metabolites were measured using the following mobile phase gradient: 0 min, 80% B; 5 min, 80% B; 30 min, 20% B; 31 min, 80% B; 42 min, 80% B. The mass spectrometer was operated in selected ion monitoring mode with an m/*z* window width of 9.0 centered at 1.003355-times half the number of carbon atoms in the target metabolite. The resolution was set at 70,000 and AGC target was 1×10^5^ ions. Peak areas were corrected for quadrupole bias as in Kim *et al.* [-@kim2015]. Raw mass isotopomer distributions were corrected for natural isotope abundance using a custom R package (mzrtools, github.com/oldhamlab/mzrtools) employing the method of Fernandez, *et al.* [-@fernandez1996].

## RNA-seq transcriptomics

@@

Research reported in this publication was supported by the National Heart Lung and Blood Institute of the National Institutes of Health under award number R01GM987654 (to W.M.O.), and by the Brigham and Women's Hospital Research Institute. The content is solely the responsibility of the authors and does not necessarily represent the official views of the National Institutes of Health. VB124 was a gift from Vettore, LLC.

>>>>

authors:
- name: David R. Ziehr
  affiliations: 
  - ref: bwh
  - ref: mgh
  - ref: hms
- name: Fei Li
  affiliations:
  - ref: bwh
  - ref: hms
- name: K. Mark Parnell
  affiliations:
  - ref: vettore
- name: Nathan M. Krah
  affiliations:
  - ref: utah_gen
- name: Kevin J. Leahy 
  affiliations:
  - ref: bwh
- name: Christelle Guillermier
  affiliations:
  - ref: bwh
  - ref: hms
- name: Jack Varon
  affiliations: 
  - ref: bwh
  - ref: hms
- name: Rebecca M. Baron
  affiliations:
  - ref: bwh
  - ref: hms
- name: Bradley A. Maron
  affiliations:
  - ref: umd
  - ref: ihc
- name: Nancy J. Philp
  affiliations:
  - ref: tj
- name: Lida P. Hariri
  affiliations: 
  - ref: hms
  - ref: mgh_path
- name: Edy Y. Kim
  affiliations:
  - ref: bwh
  - ref: hms
- name: Matthew L. Steinhauser
  affiliations:
  - ref: pitt
  - ref: upmc
- name: Rachel S. Knipe
  affiliations:
  - ref: mgh
  - ref: hms
- name: Jared Rutter
  affiliations:
  - ref: utah_biochem
  - ref: utah_hhmi
- name: William M. Oldham
  affiliations:
  - ref: bwh
  - ref: hms
  email: woldham@bwh.harvard.edu
  corresponding: true

affiliations:
- id: bwh
  name: Department of Medicine, Brigham and Women's Hospital, Boston, MA
- id: mgh
  name: Department of Medicine, Massachusetts General Hospital, Boston, MA
- id: hms
  name: Department of Medicine, Harvard Medical School, Boston, MA
- id: vettore
  name: Vettore Biosciences, San Francisco, CA
- id: utah_gen
  name: Department of Human Genetics, University of Utah, Salt Lake City, UT
- id: umd
  name: Department of Medicine, University of Maryland School of Medicine, Baltimore, MA
- id: ihc
  name: University of Maryland Institute for Health Computing, Bethesda, MD
- id: tj
  name: Department of Pathology, Anatomy & Cell Biology, Thomas Jefferson University, Philadelphia, PA
- id: mgh_path
  name: Department of Pathology, Massachusetts General Hospital, Boston, MA
- id: pitt
  name: Aging Institute, University of Pittsburgh, Pittsburgh, PA
- id: upmc
  name: UPMC Heart and Vascular Institute, UPMC Presbyterian, Pittsburgh, PA
- id: utah_biochem
  name: Department of Biochemistry, University of Utah, Salt Lake City, UT
- id: utah_hhmi
  name: Howard Hughes Medical Institute, University of Utah School of Medicine, Salt Lake City, UT

@@

Metabolites were measured using the following mobile phase gradient: 0 min, 80% B; 5 min, 80% B; 30 min, 20% B; 31 min, 80% B; 42 min, 80% B. The mass spectrometer was operated in selected ion monitoring mode with an m/*z* window width of 9.0 centered at 1.003355-times half the number of carbon atoms in the target metabolite. The resolution was set at 70,000 and AGC target was 1×10^5^ ions. Peak areas were corrected for quadrupole bias as in Kim *et al.* [-@kim2015]. Raw mass isotopomer distributions were corrected for natural isotope abundance using a custom R package (mzrtools, github.com/oldhamlab/mzrtools) employing the method of Fernandez, *et al.* [-@fernandez1996]. 
## RNA-seq transcriptomics

@@

The authors thank Louise Trakimas and the HMS Electron Microscopy Facility for MIMS sample preparation and consultation. Research reported in this publication was supported by the National Heart Lung and Blood Institute of the National Institutes of Health under award number R01GM987654 (to W.M.O.), and by the Brigham and Women's Hospital Research Institute. The content is solely the responsibility of the authors and does not necessarily represent the official views of the National Institutes of Health. VB124 was a gift from Vettore, LLC.
